\propertitle

### 2019-12-21 maxNt_v6.tex
update

\pdfoutput=1
%% Author: PGL  Porta Mana
%% Created: 2015-05-01T20:53:34+0200
%% Last-Updated: 2019-12-21T21:36:19+0100
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newif\ifarxiv
\arxivfalse
\ifarxiv\pdfmapfile{+classico.map}\fi
\newif\ifafour
\afourfalse% true = A4, false = A5
\newif\iftypodisclaim % typographical disclaim on the side
\typodisclaimtrue
\newcommand*{\memfontfamily}{zplx}
\newcommand*{\memfontpack}{newpxtext}
\documentclass[\ifafour a4paper,12pt,\else a5paper,10pt,\fi%extrafontsizes,%
onecolumn,oneside,article,%french,italian,german,swedish,latin,
british%
]{memoir}
\newcommand*{\updated}{\today}
\newcommand*{\firstdraft}{4 November 2015}
\newcommand*{\firstpublished}{***}
\newcommand*{\propertitle}{Inferring the total activity of a large neuronal population\\ from a small sample}
\newcommand*{\pdftitle}{\propertitle}
\newcommand*{\headtitle}{Inferences about large neuronal populations}
\newcommand*{\pdfauthor}{P.G.L.  Porta Mana, Y. Roudi, V. Rostami, E. Torre}
\newcommand*{\headauthor}{Porta Mana \etal}
\newcommand*{\reporthead}{}% Report number

@@

The maximum-entropy distribution for the larger population is calculated
using five moments, because this number provides almost as much information
as the full frequency distribution of the sample, as discussed in
\sect~\ref{sec:marginalization}. Figure~\ref{fig:plots_5mom} shows the
resulting distributions for three example values of
larger-population sizes: $N=1\,000$ (\textcolor{mygreen}{green diamonds}),
$N=5\,000$ (\textcolor{myred}{red circles}), $N=10\,000$
(\textcolor{myblue}{blue curve}). The frequency density of the sample
activity is also shown (\textcolor{black}{black triangles}), and in the
plot it would be indistinguishable from the maximum-entropy density for
$N=n$, that is, applied at the sample level. We discuss the case of unknown
$N$ in \sect~\ref{sec:N}.
\begin{figure}[!p]
  \centering
\subcaption[]{\label{fig:plots_5mom_a}}%  
\includegraphics[width=\linewidth]{_scripts/newpaper_stensola_5mom_distributionsN_priorc.pdf}%
\\[1em]%\mbox{}\vfill\mbox{}%
\subcaption[]{\label{fig:plots_5mom_b}}%  
\includegraphics[width=\linewidth]{_scripts/newpaper_riehle_5mom_distributionsN_priorc.pdf}%
\\[1em]%\mbox{}\vfill\mbox{}%
\caption{Maximum-entropy distributions for the total activity of the larger
  population, using five moments and assuming several population sizes $N$.
  In order to compare different $N$, the distributions are multiplied by
  $N$ (obtaining densities) and the activity divided by $N$ (obtaining a
  population average). Note how the distributions for large $N$ have very
  different features from the measured frequency distribution, and how
  their peaks become more pronounced as $N$ increases.}
\label{fig:plots_5mom}
\end{figure}
% maxent_solutions_different_moments_prior2_6moments.nb

@@

% maxent_solutions_different_moments_partial.nb
Here is a toy example of possible uses of this model, based on the first
data set. We could be interested in the hypothesis that two distinct cell
types or assemblies be present in the region where the recording was made.
Finding a larger-population distribution with two peaks, as in
\fig~\ref{fig:plots_5mom_a}, would provide some evidence for this
hypothesis. We might have reasons for suspecting that a specific set of the
sampled neurons is of the first type, and the remaining of the second type.
In the case of the first data set, 27 of the 65 sampled cells were
identified as grid cells belonging to 3--4 functional modules, the
remaining 38 \enquote{non-grid} cells being interneurons and unidentified
cells \citep{stensolaetal2012}. Could the two peaks in the distribution of
\fig~\ref{fig:plots_5mom} reflect the activities of grid versus non-grid
cells? We apply the model to these two sets of neurons individually, using
$\yNg=4\,150$ for the grid set and $\yNng=5\,850$ for the non-grid set, to
reflect their proportions (27/65 and 38/65) in the recorded sample. The
results are shown in \fig~\ref{fig:plots_cellsubsets_a}.
\begin{figure}[!p]
\centering
\subcaption[]{\label{fig:plots_cellsubsets_a}}%  
\includegraphics[width=\linewidth]{_scripts/newpaper3_gridvsnogrid.pdf}%
\\[2em]%
\subcaption[]{\label{fig:plots_cellsubsets_b}}%  
\includegraphics[width=\linewidth]{_scripts/newpaper3_gridconv.pdf}%
\\[2em]%
\caption{\subcaptionref{fig:plots_cellsubsets_a} Maximum-entropy
  distributions for larger subpopulations or grid and non-grid cells.
  \subcaptionref{fig:plots_cellsubsets_b} Convolution of the two
  subpopulation distributions; it would be equal to the total distribution
  if the two distributions were independent.}
\label{fig:plots_cellsubsets}
\end{figure}
The distribution for the larger population of grid cells
(\textcolor{mygreen}{green triangles}) seems to have one broad peak, close
to the first peak of the larger-population distribution. The distribution
for the larger population of non-grid cells (\textcolor{myred}{red
  circles}) has two peaks instead, roughly at the same population-averaged
activities as the peaks of the larger-population distribution
(\textcolor{myblue}{blue curve}) but closer in height. It would thus seem
that the population of grid cells is contributing to the first mode of the
larger population, but it is not its sole contributor. We can also assess
if the distributions for the two sets of neurons are independent. If they
were independent, the larger-population distribution would be given by
their convolution:
\begin{equation}
  \label{eq:convolution}
  P_{\text{full}}(\yA) = \sum_{\yA'} P_{\text{grid}}(\yA')\;
  P_{\text{non-grid}}(\yA-\yA') %, \qquad \yA \in \set{1,\dotsc,10\,000}
\end{equation}
where the index $\yA'$ runs from $\max(0,\yA-\yNng)$ to $\min(\yA,\yNg)$.
But this is not the case: 
figure~\ref{fig:plots_cellsubsets_b} shows that such convolution
(\textcolor{black}{black diamonds}) is quite different from the
larger-population distribution (\textcolor{myblue}{blue curve}): the two
peaks of the former are closer in position and height than those of the
latter. The population distributions of grid and non-grid cells are
therefore \emph{not} independent: knowledge of the activity of either set
gives us some information about the activity of the other.

@@

Let us illustrate the quantity just defined and the previous discussion
about sufficiency and sampling by using our data sets. Consider for example
three sets of constraints $M_{2}$, $M_{4}$, $M_{5}$, consisting of the
first two, four, five moments; and a larger population of size $N=10\,000$.
Figure~\ref{fig:plots_2_4mom} shows the resulting distributions for the two
data sets.
\begin{figure}[!p]
\centering
\subcaption[]{\label{fig:plots_2_4mom_a}}%  
\includegraphics[width=\linewidth]{_scripts/newpaper2_stensola_threemom_distributions10000_priorc.pdf}%
\\[2em]%
\subcaption[]{\label{fig:plots_2_4mom_b}}%  
\includegraphics[width=\linewidth]{_scripts/newpaper2_riehle_threemom_distributions10000_prior2.pdf}%
\\[2em]%
\caption{Maximum-entropy distributions for the normalized total activity of
  the larger population, assuming two different sets of constraining moments
  and a larger population size $N=10\,000$.}
\label{fig:plots_2_4mom}
\end{figure}
% maxent_solutions_different_moments_prior2_6moments.nb
There is a clear visual difference between the distributions based on two
(\textcolor{myred}{dashed red}) and on four (\textcolor{myblue}{solid
  blue}) moments; for the first data set, panel
\subcaptionref{fig:plots_2_4mom_a}, they even differ in the number of
modes. On the other hand there is barely any visual difference between the
distributions based on four and on five (\textcolor{myyellow}{dotted
  yellow}) moments. From these visual differences we immediately conclude
that two moments are less likely to be informationally sufficient than
four; but four should be almost as likely as five. We shall see that the
measure~\eqref{eq:measure_suffic} reflects this swift visual judgement.

@@

% NUMERICAL RESULTS
%
% relentropies of sample2mom, sample4mom, sample5mom, marg2mom, marg4mom, marg5mom
% -322.72,-42.8005,-39.4917,-85.7626,-4.23725,-4.20059 nats =
% -1401.56,-185.88,-171.51,-372.462,-18.4021,-18.2429 dB
%
% difference in relentropies between 4 mom and 2 mom:
%** sample level: 279.92 nats = 403.839 bit = 1215.68 dB
%** marg level: 81.525 nats =  117.616 bit = 354.06 dB
% difference in relentropies between 5 mom and 4 mom:
%** sample level: 3.30882 nats = 4.77362 bit = 14.37 dB
%** marg level: 0.0366623 nats = 0.0543836 bit = 0.159223 dB
% correspond to prob ratios of
% sample: 3.69524\times 10^{121}, 27.3529
% marg: 2.5469\times 10^{35}, 1.03734
% 
% comparison N vs n:
% 2 mom: 236.958 nats = 341.858 bit = 1029.09 dB, 8.11758\times 10^{102}
% 4 mom: 38.5632 nats = 55.635 bit = 167.478 dB, 5.59496*10^16
% 5 mom: 35.2911 nats = 50.9157 bit = 153.267 dB, 2.12185*10^15
%
% comparison_ME_sample_full_levels.nb
%
Calculation of the weight of evidence~\eqref{eq:measure_suffic} for the
first data set and $N=10\,000$, corresponding to
\fig~\ref{fig:plots_2_4mom_a}, leads to the following results:
\begin{equation}
  \label{eq:diff_suff_N}
  \begin{split}
  \ydi_{N}(M_{4},M_{2}) &= 118\;\bit \equiv 35\;\hart,
\\    \ydi_{N}(M_{5},M_{4}) &= 0.054\;\bit \equiv 0.016\;\hart,
  \end{split}
\end{equation}
where the Hartley ($\hart$) denotes base-10 logarithms
\citep[\sect~C.4]{iso2009}[called \enquote{ban} and used by Turing and Good
in their code-breaking work at Bletchley Park:
see][]{good1985%,good1950,good1969
}[\sect~4.2]{jaynes1994_r2003}. In words, our data are 35 orders of
magnitude more probable by assuming sufficiency of $M_{4}$ than by assuming
sufficiency of $M_{2}$. But the data are about as probable
($10^{0.016} = 1.04$) by assuming sufficiency of $M_{4}$ as of $M_{5}$. The
qualitative visual information of \fig~\ref{fig:plots_2_4mom_a} was
therefore reliable.

For the model at the sample level, that is, with $N=n$, the weights of
evidence are
\begin{equation}
  \label{eq:diff_suff_n}
  \begin{split}
  \ydi_{n}(M_{4},M_{2}) &= 400\;\bit \equiv 120\;\hart,
\\    \ydi_{n}(M_{5},M_{4}) &= 4.8\;\bit \equiv 1.4\;\hart.
\end{split}
\end{equation}
So the application at the sample level says that the data are 120 orders of
magnitude more probable conditional on $M_{4}$ than on $M_{2}$, and
$10^{1.4}=25$ times more probable conditional on $M_{5}$ than on $M_{4}$.
These results are different from those of the application at the larger
population, and they would be hard to deduce visually from the
distributions of \fig~\ref{fig:comparisons_marginals}.

% The conclusions about \enquote{cooperativity} or \enquote{interaction} that
% we reach by applying the maximum-entropy method to the larger population,
% as proposed here, are therefore different from those applying it at the
% sample level -- and also visually clearer. The plot in 
% \fig~\ref{fig:plots_2_4mom_a}, showing the two clearly different
% distributions constrained by two and four moments, should be compared with
% the plot for the distributions obtained at the sample level under the same
% constraints, shown in \fig~\ref{fig:comparisons_marginals} as
% \textcolor{myred}{red circles} and \textcolor{myblue}{blue squares}. The
% measured frequencies (black triangles) are also shown. It is necessary to
% use a logarithmic scale to see the differences, which appear mainly in the
% tail.
\begin{figure}[!p]
\centering
% \includegraphics[width=\linewidth]{_scripts/comparison_full.pdf}%
% \\[3em]%
\includegraphics[width=0.95\linewidth]{_scripts/trunct2_comparison_marginals.pdf}%
\caption{Marginals for the normalized total activity of the sample,
  obtained from the maximum-entropy distributions with $N=10\,000$ from
  \fig~\ref{fig:plots_2_4mom_a}, and from maximum-entropy distributions
  calculated at the sample level, that is with $N=n=65$, with two and four
  moments in either case.}
\label{fig:comparisons_marginals}
\end{figure}
% comparison_ME_sample_full_levels.nb

@@

It is possible to use the measure~\eqref{eq:measure_suffic} and its
probabilistic interpretation~\eqref{eq:bayes_factor} to compare the
probability of the data $\yf$ conditional on the hypothesis $(M_{5},N)$ of
four-moment sufficiency at the larger-population level, $N=10\,000$, versus
the hypothesis $(M_{5}, n)$ of four-moment sufficiency at the sample level,
$N=n=65$. We obtain
\begin{equation}
  \label{eq:diff_suff_n}
  \ydi[(M_{5}, N), (M_{5}, n)] = 51\;\bit \equiv 15\;\hart.
\end{equation}
% that is, the data are 15 orders of magnitude more probable under the first
% hypothesis than under the second. 
This result is also shown in
\fig~\ref{fig:comparisons_marginals}. We see that the four-moment
distribution constructed at the larger-population level
(\textcolor{myblue}{blue filled squares}) fits the measured frequency
distribution (black triangles) more closely than the corresponding
sample-level distribution (\textcolor{myblue}{blue empty squares}).

>>>>

\pdfoutput=1
%% Author: PGL  Porta Mana
%% Created: 2015-05-01T20:53:34+0200
%% Last-Updated: 2019-12-21T23:54:35+0100
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newif\ifarxiv
\arxivfalse
\ifarxiv\pdfmapfile{+classico.map}\fi
\newif\ifafour
\afourfalse% true = A4, false = A5
\newif\iftypodisclaim % typographical disclaim on the side
\typodisclaimtrue
\newcommand*{\memfontfamily}{zplx}
\newcommand*{\memfontpack}{newpxtext}
\documentclass[\ifafour a4paper,12pt,\else a5paper,10pt,\fi%extrafontsizes,%
onecolumn,oneside,article,%french,italian,german,swedish,latin,
british%
]{memoir}
\newcommand*{\updated}{\today}
\newcommand*{\firstdraft}{4 November 2015}
\newcommand*{\firstpublished}{***}
\newcommand*{\propertitle}{Inferring the total activity of a large neuronal population\\ from a small sample}
\newcommand*{\pdftitle}{\propertitle}
\newcommand*{\headtitle}{Inferences about large neuronal populations}
\newcommand*{\pdfauthor}{P.G.L.  Porta Mana, Y. Roudi, V. Rostami, E. Torre}
\newcommand*{\headauthor}{Porta Mana \etal}
\newcommand*{\reporthead}{}% Report number

@@

The maximum-entropy distribution for the larger population is calculated
using five moments, because this number provides almost as much information
as the full frequency distribution of the sample, as discussed in
\sect~\ref{sec:marginalization}. Figure~\ref{fig:plots_5mom} shows the
resulting distributions for three example values of
larger-population sizes: $N=1\,000$ (\textcolor{mygreen}{green diamonds}),
$N=5\,000$ (\textcolor{myred}{red circles}), $N=10\,000$
(\textcolor{myblue}{blue curve}). The frequency density of the sample
activity is also shown (\textcolor{black}{black triangles}), and in the
plot it would be indistinguishable from the maximum-entropy density for
$N=n$, that is, applied at the sample level. We discuss the case of unknown
$N$ in \sect~\ref{sec:N}.
\begin{figure}[!p]
  \centering
\subcaption[]{\label{fig:plots_5mom_a}}%  
\includegraphics[width=\linewidth]{_scripts/newpaper_stensola_5mom_distributionsN_priorc.pdf}%
\\[1em]%\mbox{}\vfill\mbox{}%
\subcaption[]{\label{fig:plots_5mom_b}}%  
\includegraphics[width=\linewidth]{_scripts/newpaper_riehle_5mom_distributionsN_priorc.pdf}%
\\[1em]%\mbox{}\vfill\mbox{}%
\caption{Maximum-entropy distributions for the total activity of the larger
  population, using five moments and assuming several population sizes $N$.
  In order to compare different $N$, the distributions are multiplied by
  $N$ (obtaining densities) and the activity divided by $N$ (obtaining a
  population average). The distributions for large $N$ have very different
  features from the measured frequency distribution, and how their peaks
  become more pronounced as $N$ increases.}
\label{fig:plots_5mom}
\end{figure}
% maxent_solutions_different_moments_prior2_6moments.nb

@@

% maxent_solutions_different_moments_partial.nb
Here is a toy example of possible uses of this model, based on the first
data set. We could be interested in the hypothesis that two distinct cell
types or assemblies be present in the region where the recording was made.
Finding a larger-population distribution with two peaks, as in
\fig~\ref{fig:plots_5mom_a}, would provide some evidence for this
hypothesis. We might have reasons for suspecting that a specific set of the
sampled neurons is of the first type, and the remaining of the second type.
In the case of the first data set, 27 of the 65 sampled cells were
identified as grid cells belonging to 3--4 functional modules, the
remaining 38 \enquote{non-grid} cells being interneurons and unidentified
cells \citep{stensolaetal2012}. Could the two peaks in the distribution of
\fig~\ref{fig:plots_5mom} reflect the activities of grid versus non-grid
cells? We apply the model to these two sets of neurons individually, using
$\yNg=4\,150$ for the grid set and $\yNng=5\,850$ for the non-grid set, to
reflect their proportions (27/65 and 38/65) in the recorded sample. The
results are shown in \fig~\ref{fig:plots_cellsubsets_a}.
\begin{figure}[!p]
\centering
\subcaption[]{\label{fig:plots_cellsubsets_a}}%  
\includegraphics[width=\linewidth]{_scripts/newpaper3_gridvsnogrid.pdf}%
\\[1em]%
\subcaption[]{\label{fig:plots_cellsubsets_b}}%  
\includegraphics[width=\linewidth]{_scripts/newpaper3_gridconv.pdf}%
\\[1em]%
\caption{\subcaptionref{fig:plots_cellsubsets_a} Maximum-entropy
  distributions for larger subpopulations or grid and non-grid cells.
  \subcaptionref{fig:plots_cellsubsets_b} The convolution of the two
  subpopulation distributions is very different from the distribution for
  all cells; this means that the two subpopulation distributions are not
  independent.}
\label{fig:plots_cellsubsets}
\end{figure}
The distribution for the larger population of grid cells
(\textcolor{mygreen}{green triangles}) seems to have one broad peak, close
to the first peak of the larger-population distribution. The distribution
for the larger population of non-grid cells (\textcolor{myred}{red
  circles}) has two peaks instead, roughly at the same population-averaged
activities as the peaks of the larger-population distribution
(\textcolor{myblue}{blue curve}) but closer in height. It would thus seem
that the population of grid cells is contributing to the first mode of the
larger population, but it is not its sole contributor. We can also assess
if the distributions for the two sets of neurons are independent. If they
were independent, the larger-population distribution would be given by
their convolution:
\begin{equation}
  \label{eq:convolution}
  P_{\text{full}}(\yA) = \sum_{\yA'} P_{\text{grid}}(\yA')\;
  P_{\text{non-grid}}(\yA-\yA') %, \qquad \yA \in \set{1,\dotsc,10\,000}
\end{equation}
where the index $\yA'$ runs from $\max(0,\yA-\yNng)$ to $\min(\yA,\yNg)$.
But this is not the case: 
figure~\ref{fig:plots_cellsubsets_b} shows that such convolution
(\textcolor{black}{black diamonds}) is quite different from the
larger-population distribution (\textcolor{myblue}{blue curve}): the two
peaks of the former are closer in position and height than those of the
latter. The population distributions of grid and non-grid cells are
therefore \emph{not} independent: knowledge of the activity of either set
gives us some information about the activity of the other.

@@

Let us illustrate the quantity just defined and the previous discussion
about sufficiency and sampling by using our data sets. Consider for example
three sets of constraints $M_{2}$, $M_{4}$, $M_{5}$, consisting of the
first two, four, five moments; and a larger population of size $N=10\,000$.
Figure~\ref{fig:plots_2_4mom} shows the resulting distributions for the two
data sets.
\begin{figure}[!p]
\centering
\subcaption[]{\label{fig:plots_2_4mom_a}}%  
\includegraphics[width=\linewidth]{_scripts/newpaper2_stensola_threemom_distributions10000_priorc.pdf}%
\\[1em]%
\subcaption[]{\label{fig:plots_2_4mom_b}}%  
\includegraphics[width=\linewidth]{_scripts/newpaper2_riehle_threemom_distributions10000_prior2.pdf}%
\\[1em]%
\caption{Maximum-entropy distributions for the total activity of the larger
  population, based on the first two, four, and five moments, and a
  larger-population size $N=10\,000$. The two- and four-moment
  distributions are very different, suggesting that two moments are not
  informationally sufficient. The four- and five-moment distributions are
  almost indistinguishable, suggesting that four and five moments are
  equally likely to be sufficient.}
\label{fig:plots_2_4mom}
\end{figure}
% maxent_solutions_different_moments_prior2_6moments.nb
There is a clear visual difference between the distributions based on two
(\textcolor{myred}{dashed red}) and on four (\textcolor{myblue}{solid
  blue}) moments; for the first data set, panel
\subcaptionref{fig:plots_2_4mom_a}, they even differ in the number of
modes. On the other hand there is barely any visual difference between the
distributions based on four and on five (\textcolor{myyellow}{dotted
  yellow}) moments. From these visual differences we immediately conclude
that two moments are less likely to be informationally sufficient than
four; but four should be almost as likely as five. We shall see that the
measure~\eqref{eq:measure_suffic} reflects this swift visual judgement.

@@

% NUMERICAL RESULTS
%
% relentropies of sample2mom, sample4mom, sample5mom, marg2mom, marg4mom, marg5mom
% -322.72,-42.8005,-39.4917,-85.7626,-4.23725,-4.20059 nats =
% -1401.56,-185.88,-171.51,-372.462,-18.4021,-18.2429 dB
%
% difference in relentropies between 4 mom and 2 mom:
%** sample level: 279.92 nats = 403.839 bit = 1215.68 dB
%** marg level: 81.525 nats =  117.616 bit = 354.06 dB
% difference in relentropies between 5 mom and 4 mom:
%** sample level: 3.30882 nats = 4.77362 bit = 14.37 dB
%** marg level: 0.0366623 nats = 0.0543836 bit = 0.159223 dB
% correspond to prob ratios of
% sample: 3.69524\times 10^{121}, 27.3529
% marg: 2.5469\times 10^{35}, 1.03734
% 
% comparison N vs n:
% 2 mom: 236.958 nats = 341.858 bit = 1029.09 dB, 8.11758\times 10^{102}
% 4 mom: 38.5632 nats = 55.635 bit = 167.478 dB, 5.59496*10^16
% 5 mom: 35.2911 nats = 50.9157 bit = 153.267 dB, 2.12185*10^15
%
% comparison_ME_sample_full_levels.nb
%
Calculation of the weight of evidence~\eqref{eq:measure_suffic} for the
first data set and $N=10\,000$, corresponding to
\fig~\ref{fig:plots_2_4mom_a}, leads to the following results:
\begin{equation}
  \label{eq:diff_suff_N}
  \begin{split}
  \ydi_{N}(M_{4},M_{2}) &= 118\;\bit \equiv 35.4\;\hart,
\\    \ydi_{N}(M_{5},M_{4}) &= 0.0544\;\bit \equiv 0.0159\;\hart,
  \end{split}
\end{equation}
where the Hartley ($\hart$) denotes base-10 logarithms
\citep[\sect~C.4]{iso2009}[called \enquote{ban} and used by Turing and Good
in their code-breaking work at Bletchley Park:
see][]{good1985%,good1950,good1969
}[\sect~4.2]{jaynes1994_r2003}. In words, our data are 35 orders of
magnitude more probable by assuming sufficiency of $M_{4}$ than by assuming
sufficiency of $M_{2}$. But the data are about as probable
($10^{0.016} = 1.04$) by assuming sufficiency of $M_{4}$ as of $M_{5}$. The
qualitative visual information of \fig~\ref{fig:plots_2_4mom_a} was
therefore reliable.

For the model at the sample level, that is, with $N=n$, the weights of
evidence are
\begin{equation}
  \label{eq:diff_suff_n}
  \begin{split}
  \ydi_{n}(M_{4},M_{2}) &= 404\;\bit \equiv 122\;\hart,
\\    \ydi_{n}(M_{5},M_{4}) &= 4.77\;\bit \equiv 1.44\;\hart.
\end{split}
\end{equation}
So the application at the sample level says that the data are 120 orders of
magnitude more probable conditional on $M_{4}$ than on $M_{2}$, and
$10^{1.4}=25$ times more probable conditional on $M_{5}$ than on $M_{4}$.
These results are different from those of the application at the larger
population, and they would be hard to deduce visually from the
distributions of \fig~\ref{fig:comparisons_marginals}.

% The conclusions about \enquote{cooperativity} or \enquote{interaction} that
% we reach by applying the maximum-entropy method to the larger population,
% as proposed here, are therefore different from those applying it at the
% sample level -- and also visually clearer. The plot in 
% \fig~\ref{fig:plots_2_4mom_a}, showing the two clearly different
% distributions constrained by two and four moments, should be compared with
% the plot for the distributions obtained at the sample level under the same
% constraints, shown in \fig~\ref{fig:comparisons_marginals} as
% \textcolor{myred}{red circles} and \textcolor{myblue}{blue squares}. The
% measured frequencies (black triangles) are also shown. It is necessary to
% use a logarithmic scale to see the differences, which appear mainly in the
% tail.
\begin{figure}[!p]
\centering
% \includegraphics[width=\linewidth]{_scripts/comparison_full.pdf}%
% \\[3em]%
\includegraphics[width=0.95\linewidth]{_scripts/trunct2_comparison_marginals.pdf}%
\caption{Maximum-entropy distributions constructed at the sample level
  ($N=n$), based on the first two, four, and five moments; and sample
  marginal of the distribution constructed with five moments for a larger
  population with $N=10\,000$ (dotted yellow in
  \subcaptionref{fig:plots_2_4mom_a}). The distributions at the sample
  level are all very similar, only distinguishable on a logarithmic scale.
  The five-moment distribution constructed at the larger-population level
  fits the frequency data better than the one at the sample level.}
\label{fig:comparisons_marginals}
\end{figure}
% comparison_ME_sample_full_levels.nb

@@

It is possible to use the measure~\eqref{eq:measure_suffic} and its
probabilistic interpretation~\eqref{eq:bayes_factor} to compare the
probability of the data $\yf$ conditional on the hypothesis $(M_{5},N)$ of
five-moment sufficiency at the larger-population level, $N=10\,000$, versus
the hypothesis $(M_{5}, n)$ of five-moment sufficiency at the sample level,
$N=n=65$. We obtain
\begin{equation}
  \label{eq:diff_suff_n}
  \ydi[(M_{5}, N), (M_{5}, n)] = 50.9\;\bit \equiv 15.3\;\hart.
\end{equation}
% that is, the data are 15 orders of magnitude more probable under the first
% hypothesis than under the second. 
This result is also shown in
\fig~\ref{fig:comparisons_marginals}. We see that the five-moment
distribution constructed at the larger-population level
(\textcolor{myyellow}{yellow stars}) fits the measured frequency
distribution (black triangles) more closely than the corresponding
sample-level distribution (\textcolor{myyellow}{yellow lozenges}).
